---
title: "ema_grant_models.rmd"
author: "matthew"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

If Question 14 is tagged as positive (negative event occurred recently), we would use question 15 as an index of the negativity (a covariate), an average of 17 and 18 as a marker of focus at time of event (potential covariate unless colinear with predictor), and average of questions 5-12 (rumination habit) to predict an average of questions 1- 4 (INVERT 2)

```{r}
library(tidyverse)
library(lme4)
library(performance)
library(emmeans)
library(modelbased)
```

askd

```{r}
df = read_csv("baseline_r_20240222.csv")
df = df %>%
  rename(
    response_id=Response_Id,
    user_id=User_Id,
    start_date=Survey_Started_Date,
    start_time=Survey_Started_Time,
    submit_date=Survey_Submitted_Date,
    submit_time=Survey_Submitted_Time,
    time_zone=Time_Zone,
    study_day=Study_Day,
    feeling_sad=Start_Feeling_Sad,
    feeling_happy=Start_Feeling_Happy,
    feeling_anxious=Start_Feeling_anxious,
    feeling_annoyed=Start_Feeling_Annoyed,
    rum_1=Start_Feeling_Handle,
    rum_2=Start_Feeling_Problems,
    rum_3=Start_Feeling_move_forward,
    rum_4=Start_Feeling_Sensations,
    rum_5=Start_Feeling_Again,
    rum_6=Start_Feeling_Wanting,
    rum_7=Start_Feeling_Stuck,
    rum_8=Start_Feeling_Negative_Thoughts,
    event_occured=quest_1676672138796,
    event_mood=quest_1677010062008,
    event_arousal=Time_Of_Event_Calm,
    event_focus_self=Why_I_Felt,
    event_focus_env=Why_I_Felt_1658258003578
    ) %>%
  select(
    response_id,user_id,start_date,start_time,submit_date,submit_time,time_zone,
    study_day,feeling_sad,feeling_happy,feeling_anxious,feeling_annoyed,
    rum_1,rum_2,rum_3,rum_4,rum_5,rum_6,rum_7,rum_8,event_occured,event_mood,
    event_arousal,event_focus_self,event_focus_env
  ) %>%
  mutate_all(~ replace(.,.=="CONDITION_SKIPPED",NA)) %>%
  mutate(
    event_mood = as.double(event_mood),
    event_arousal = as.double(event_arousal),
    event_focus_self = as.double(event_focus_self),
    event_focus_env = as.double(event_focus_env)) %>%
  rowwise() %>%
  mutate(
    feeling_happy_rev = 6 - feeling_happy
  ) %>%
  mutate(
    mood_avg = mean(c(feeling_sad,feeling_happy_rev,feeling_anxious,feeling_annoyed)),
    rum_avg = mean(c(rum_1,rum_2,rum_3,rum_4,rum_5,rum_6,rum_7,rum_8)),
    event_focus_avg = mean(c(event_focus_self,event_focus_env))) %>%
  ungroup()
df
```

negative mood: 1-4
rumination: 5-12
negative event happened: 14
mood at event: 15
focus at event: 17,18

Model:

Negative mood (average) ~ Rumination (average) + Negative event happened recently + Mood at time of event + Focus at time of event

```{r}
model = lmer(
  mood_avg ~ rum_avg + event_occured + event_mood + event_focus_avg + (1 + rum_avg|user_id), 
  data = df,
  REML = F)

model_null = lmer(
  mood_avg ~ event_occured + event_mood + event_focus_avg + (1 + rum_avg|user_id), 
  data = df,
  REML = F)

summary(model)
```

```{r}
anova(model_null,model)
```

```{r}
rum.emm = emmeans(model,~rum_avg)
rum.emm
```

```{r}
check_collinearity(model)
```

```{r,fig.height=12,fig.width=15}
check_model(model)
```

```{r}
check_normality(model)
```
