---
title: "ema_baseline"
author: "matthew"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


Our objective is to determine associations between mood and rumination
The relevant variables here seem to be ratings of happiness & sadness, and the 8 rumination items

Let's try making 2 models: sad ~ rumination sum, happy ~ rumination sum

But first explore the data a bit


```{r,quiet=TRUE}
library(tidyverse)
library(ggstance)
library(rmcorr)
library(lme4)
library(ordinal)
```


Load the data

```{r}
df = read_csv('baseline_r_20240222.csv')
head(df)
```


total number of submissions
```{r}
df %>%
  select(Response_Id) %>%
  unique() %>%
  nrow()
```


number of particpants
```{r}
df %>%
  select(User_Id) %>%
  unique() %>%
  nrow()
```


submissions per participant
```{r}
df %>%
  group_by(User_Id) %>%
  count() %>%
  summary()
```


survey submission time distribution
```{r}

# df %>%
#   group_by(Response_Id) %>%
#   select(c(Response_Id,Survey_Submitted_Time)) %>%
#   #mutate(Survey_Submitted_Time = as.POSIXlt(Survey_Submitted_Time)) %>%
#   #mutate(Survey_Submitted_Time = as.POSIXct(Survey_Submitted_Time)) %>%
#   ggplot(aes(x=1:nrow(df),y=Survey_Submitted_Time)) +
#   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.5)
#   #scale_y_continuous(trans = rev_date)
#   #scale_y_reverse()

```


calculate rumination score
sum of all rumination items (5-12)
```{r}

df = df %>%
  mutate(User_Id = as.factor(User_Id)) %>%
  mutate(rumination = rowSums(select(.,c(
    Start_Feeling_Handle,Start_Feeling_Problems,Start_Feeling_move_forward,
    Start_Feeling_Sensations,Start_Feeling_Again,Start_Feeling_Wanting,
    Start_Feeling_Stuck,Start_Feeling_Negative_Thoughts))
    ))

```


scatter plot: sad vs. rum
fit a regression line for each participant
```{r}
df %>%
  ggplot(show.legend=F) +
  geom_point(aes(x=rumination,y=Start_Feeling_Sad,color=User_Id),show.legend=F,
             position=ggstance::position_dodgev(height=0.3)) +
  geom_smooth(aes(x=rumination,y=Start_Feeling_Sad,color=User_Id),method="lm",se=F,
              show.legend=F) +
  geom_smooth(aes(x=rumination,y=Start_Feeling_Sad),method="lm",se=F,
              show.legend=F,color="black",linewidth=1.5,linetype=2) +
  labs(title = "Negative mood vs. Rumination", subtitle = "(colored by participant)",
       x="Rumination (sum of rumination items)", y="Sadness Rating (1-5)")
```


Let's separate the plots by participant
```{r,fig.width=12,fig.height=10}
df %>%
  ggplot(aes(x=rumination,y=Start_Feeling_Sad,color=User_Id),show.legend=F) +
  geom_point(show.legend=F,position=ggstance::position_dodgev(height=0.3)) +
  geom_smooth(method = "lm",se = F, show.legend=F) +
  facet_wrap(~User_Id) +
  labs(title = "Negative mood vs. Rumination", subtitle = "(colored by participant)",
       x="Rumination (sum of rumination items)", y="Sad Rating (1-5)")
```


Same for positive mood
scatter plot: happy vs. rum
```{r}
df %>%
  ggplot(show.legend=F) +
  geom_point(aes(x=rumination,y=Start_Feeling_Happy,color=User_Id),show.legend=F,
             position=ggstance::position_dodgev(height=0.3)) +
  geom_smooth(aes(x=rumination,y=Start_Feeling_Happy,color=User_Id),method="lm",se=F,
              show.legend=F) +
  geom_smooth(aes(x=rumination,y=Start_Feeling_Happy),method="lm",se=F,
              show.legend=F,color="black",linewidth=1.5,linetype=2) +
  labs(title = "Positive mood vs. Rumination", subtitle = "(colored by participant)",
       x="Rumination (sum of rumination items)", y="Happy Rating (1-5)")
```


```{r,fig.width=12,fig.height=10}
df %>%
  ggplot(aes(x=rumination,y=Start_Feeling_Happy,color=User_Id),show.legend=F) +
  geom_point(show.legend=F,position=ggstance::position_dodgev(height=0.3)) +
  geom_smooth(method = "lm",se=F,show.legend=F) +
  facet_wrap(~User_Id) +
  labs(title = "Positive mood vs. Rumination", subtitle = "(colored by participant)",
       x="Rumination (sum of rumination items)", y="Happy Rating (1-5)")
```


*rmcorr*
Let's try repeated measures correlation to evaluate the association between
mood and rumination

First, negative mood (sad) correlated with rumination
```{r}
scorr = rmcorr(
  participant = User_Id, 
  measure1 = rumination, 
  measure2 = Start_Feeling_Sad,
  dataset = df)

scorr
```


A not very useful plot of the model
Note that rmcorr allows subjects to have different intercepts, but not slopes
```{r}
splt_label <- paste0("r = ", round(scorr$r, 3), " (95% CI: [", paste0(round(scorr$CI, 3), collapse = ", "), "])")
#par(mar = rep(2, 4))
plot(scorr, main = splt_label, xlab = "Rumination", ylab = "Sad rating")
```


Now the same for positive mood
```{r}
hcorr = rmcorr(
  participant = User_Id, 
  measure1 = rumination, 
  measure2 = Start_Feeling_Happy,
  dataset = df)

hcorr
```


```{r}
hplt_label <- paste0("r = ", round(hcorr$r, 3), " (95% CI: [", paste0(round(hcorr$CI, 3), collapse = ", "), "])")
#par(mar = rep(2, 4))
plot(hcorr, main = hplt_label, xlab = "Rumination", ylab = "Happy rating")
```



*lmer*

Now let's try some mixed modelling
LMM effect of rumination score on sadness rating
add random intercept & slope for participant
test significance via likelihood ratio
```{r}
lms_n = lmer(Start_Feeling_Sad ~ 1 + (1 + rumination|User_Id), data = df, REML = FALSE)
lms = lmer(Start_Feeling_Sad ~ rumination + (1 + rumination|User_Id), data = df, REML = FALSE)
summary(lms)
```

```{r}
anova(lms,lms_n)
```

```{r}
lmh_n = lmer(Start_Feeling_Happy ~ 1 + (1 + rumination|User_Id), data = df, REML = FALSE)
lmh = lmer(Start_Feeling_Happy ~ rumination + (1 + rumination|User_Id), data = df, REML = FALSE)
summary(lmh)
```

```{r}
anova(lmh,lmh_n)
```

According to these models, there is a significant positive effect of rumination on negative mood
& negative effect of rumination on positive mood


BUT!
Likert data are ordinal and violate the assumptions of GLM
(linear mixed models may or may not be appropriate for ordinal data)

to properly model ordinal data, let's use cumulative linked mixed models
from the ordinal r package

first, change dependent variable datatypes to ordinal
```{r}
df_o = df %>%
  mutate(Start_Feeling_Sad = as.ordered(Start_Feeling_Sad),
         Start_Feeling_Happy = as.ordered(Start_Feeling_Happy))
df_o
```

model builds similar to lme4
```{r}
clms_n = clmm(Start_Feeling_Sad ~ 1 + (1 + rumination|User_Id), data = df_o)
clms = clmm(Start_Feeling_Sad ~ rumination + (1 + rumination|User_Id), data = df_o)
clms
```

```{r}
anova(clms,clms_n)
```

Per AIC, there is less unexplained variation when rumination is included in
the model compared to when it is not

The chisquared value is significant

Now the same CLM for positive mood
```{r}
clmh_n = clmm(Start_Feeling_Happy ~ 1 + (1 + rumination|User_Id), data = df_o)
clmh = clmm(Start_Feeling_Happy ~ rumination + (1 + rumination|User_Id), data = df_o)
clmh
```

```{r}
anova(clmh_n,clmh)
```

All of the approaches show a momentary positive relationship between rumination and negative mood,
and a momentary negative relationship between rumination and positive mood



```{r}



```




